---
title             : "Replication in Social Psychology"
shorttitle        : "Replication Social Psychology"
author: 
  - name          : "Jordan Mark Barbone"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "jb769904@wcupa.edu"
affiliation:
  - id            : "1"
    institution   : "West Chester University of Pennsylvania University"
authornote: |
  Department of Psychology, 125 West Rosedale Avenue, West Chester University of Pennsylvania, West Chester, PA, 19383.
  Paper prepared for PSY609 Advanced Social Psychology.
abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
  
  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
  
  One sentence clearly stating the **general problem** being addressed by  this particular study.
  
  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
  
  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
  
  One or two sentences to put the results into a more **general context**.
  
  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
  
  
  <!-- https://tinyurl.com/ybremelq -->
keywords          : "Social psychology, Replication, Collaboration"
wordcount         : "X"
bibliography      : references.bib
floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
documentclass     : "apa6"
classoption       : "man"
#output            : papaja::apa6_pdf
output            : papaja::apa6_docx
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
library("papaja")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

This is the introcuction to my paper.  it is very well written.  So here's a citation [@open2012open] or something and [@open2015estimating].  Oh, here's another reference [@klein2014investigating] and this ought to be important.

# Methods
We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

## Participants

No participants were used here.  This is a literature review.

## Material

Well, yeah no.

## Procedure

I searched GoogeScholar.

## Data analysis

None.

# Results

We're going to have a Conclusions section I believe.

# Discussion

See above.

\newpage

# References
```{r create_r-references}
r_refs(file = "rreferences.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup
